The Astropy Project: Building an open-science project and status of the v2.0 core package

### 2018-03-01 main.tex

\begin{abstract}
The \astropy Project supports and fosters the development of open-source and openly-developed
\python packages that provide commonly-needed functionality to the astronomical
community.
A key element of the \astropy Project is the core package \astropypkg, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate and support
development for a broader ecosystem of inter-operable packages.
We conclude with a future outlook of planned new features and directions for the
broader \astropy Project.
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
All modern astronomical research makes use of software in some way.
Astronomy as a field has thus long supported the development of software tools
for astronomical tasks, such as scripts that enable individual scientific
research, software packages for small collaborations, and data reduction
pipelines for survey operations.
Some software packages are, or were, supported by large institutions and are
intended for a wide range of users.
These packages therefore typically provide some level of documentation and user
support or training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
For both packages meant for wider distribution and for scripts specific to
particular research projects, a library that addresses common astronomical tasks
simplifies the software development process.
The users of such a library then also benefit from a community and ecosystem
built around a shared foundation.
The \astropy Project has grown to become this community for \python astronomy
software, and the \astropypkg core package is a feature-rich \python library.

@@

The \astropy Project aims to provide an open-source and open-development core
package (\astropypkg) and an ecosystem of \emph{affiliated packages} that
support astronomical functionality in the \python programming language.
The \astropypkg core package is now a feature-rich library of sufficiently
general tools and classes that supports the development of more specialized
code. An example of such functionality is reading and writing FITS files: It would be
time consuming and impractical for multiple groups to implement the FITS
standard \citep{FITS} and maintain software for such a general-purpose need.
Another example of such a common task is in dealing with representations of and
transformations between astronomical coordinate systems.

The \astropy Project aims to develop and provide high-quality code and
documentation according to the best practices in software development.
The project makes use of different tools and web services to reach those
goals without central institutional oversight.
The first public release of the \astropypkg package is described in
\cite{astropy}. Since then, the \astropypkg package has been
used in hundreds of projects and the scope of the package has grown
considerably. At the same time, the scientific community
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropypkg core has
developed.
In this paper, we describe the current status of the \astropy community and the
\astropypkg core package and discuss goals for future development.

We start by describing the way the \astropy Project functions and is organized
in \sectionname~\ref{sec:org}. We then describe the main software efforts
developed by the \astropy Project itself: a core package called \astropypkg
(\sectionname~\ref{sec:core}) and several separate packages that help maintain
the infrastructure for testing and documentation
(\sectionname~\ref{sec:infrastructure}). We end with a short vision for
the future of \astropy and astronomical software in general in
\sectionname~\ref{sec:future}. The full paper, including the code
to produce the figures, is available in a \github
repository\footnote{\url{https://github.com/astropy/astropy-v2.0-paper}}

@@

A major part of the \astropy Project is the concept of ``Affiliated Packages.''
An affiliated package is an astronomy-related
\python package that is not part of the \astropypkg core package, but
has requested to be included as part of the \astropy Project's
community. These packages support the goals and vision of \astropy of
improving code re-use, interoperability, and embracing good coding
practices such as testing and thorough documentation.

Affiliated packages contain functionality that is more specialized,
have license incompatibilities, or have external dependencies (e.g., GUI
libraries) that make these packages more suitable to be separate from the
\astropypkg core package.
Affiliated packages may also be used to develop substantial new functionality
that will eventually be incorporated into the \astropypkg core package
(e.g., \texttt{wcsaxes}).
New functionality benefits from having a rapid development and release cycle that is not tied to that of the \astropypkg core (\sectionname~\ref{sect:releasecycle}).

Affiliated packages are listed on the main \astropy website and advertised to
the community through \astropy mailing lists; a list of current affiliated
packages is included in \tablename~\ref{tab:registry}.
Becoming an affiliated package is a good way for new and existing packages to
gain exposure while promoting \astropy's high standard for code and
documentation quality.
This process of listing and promoting affiliated packages is one way in which
the \astropy Project tries to increase code re-use in the astronomical
community.

@@

The \astropy Project, as of the version 2.0 release, does not receive any direct
financial support for the development of \astropypkg.
Development of the software, all supporting materials, and community support are
provided by individuals who work on the \astropy Project in their own personal
time, by individuals or groups contributing to \astropy as part of a research
project, or contributions from institutions that allocate people to work on
\astropy.
A list of organizations that have contributed to \astropy in this manner
can be found in the Acknowledgments.

Different funding models have been proposed for support of \astropy
(e.g., \citealt{2016arXiv161003159M}), but a long-term plan
for sustainability has not yet been established.
The \astropy Project has the ability to accept financial contributions
from institutions or individuals through the NumFOCUS\footnote{NumFOCUS
is a 501(c)(3) nonprofit that supports and promotes world-class, innovative,
open source scientific computing.} organization. NumFOCUS has, to date, covered the direct costs incurred by the \astropy Project.

\subsection{Reuse of the scientific Python ecosystem}

The \astropy Project is built on a philosophy of building from the existing
Python scientific ecosystem wherever possible. Many of the enabling technologies
like \texttt{numpy}, \texttt{scipy}, \texttt{matplotlib}, or \texttt{cython},
are necessary as the core underpinnings of \astropy packages.  Often this means
using these packages as ``building blocks'' (e.g., the ubiquitous
use of \texttt{numpy} arrays throughout \astropypkg).  In other cases this means
wrappers around more general algorithms that make them more convenient for
astronomy use cases (e.g., the model-fitting in the \package{astropy.modeling}
subpackage discussed in \sectionname~\ref{sec:modeling}).  Sometimes these
simple wrappers evolve into more complex implementations that address
astronomically-relevant use cases the general tool does not support (e.g.,
\package{astropy.convolution}, see \sectionname~\ref{ssec:convolution}). As a
broad rule, the Project explicitly encourages re-use of code where possible.

However, the boundaries of when this re-use is called for often ambiguous.  Some
of the examples above only evolved after significant debate in the community
over whether these algorithms were sufficient.  Other times, even apparently
general functionality did not exist in the wider ecosystem that met the \astropy
community's needs, and hence the functionality had to be developed wholly from
the developer resources available in the community (e.g.,
\package{astropy.units}, or \package{astropy.table} when it began). While some
of this functionality is ``given back'' to the wider community (e.g., myriad bug
fixes for the enabling technologies listed above, and some of the testing and
documentation architecture), in general, careful consideration is taken of
whether the additional effort to make a piece of \astropy, and it is only
``extracted'' when it seems plausible that the ongoing effort is worth the
reward in the form of additional contributors or maintainers.

\section{Astropy Core Package Version 2.0}
\label{sec:core}
The \astropy Project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg core package, which provides general
functionality (e.g., coordinate transformations, reading and writing
astronomical files, and units) or base classes for other
packages to utilize for a common interface (e.g., \texttt{NDData}).
In this section, we highlight new features introduced or substantially improved
since version 0.2 (previously described in \citealt{astropy}).  The \astropypkg
package
provides a full log of changes\footnote{\url{https://github.com/astropy/astropy/blob/stable/CHANGES.rst}}
over the course of the entire project and more details about individual
subpackages are available in the documentation.\footnote{\url{http://docs.astropy.org/en/stable/}}
Beyond what is mentioned below, most subpackages have seen improved performance
since the release of the version 0.2 package.

@@

In order to check the accuracy of the coordinate transformations in
\package{astropy.coordinates}, we have created a set of benchmarks that we use
to compare transformations between a set of coordinate frames for a number of
packages\footnote{\url{http://www.astropy.org/coordinates-benchmark/summary.html}}.
Since no package can be guaranteed to implement all transformations to
arbitrary precision and some transformations are sometimes subject to
interpretation of standards (in particular in the case of Galactic coordinates),
we do not designate any of the existing packages as the ``ground truth'' but
instead compare each tool to all other tools. The benchmarks are thus useful
beyond the \astropy Project since they allow all of the tools to be compared to
all other tools. The tools included in the benchmark at the moment include the
\astropypkg core package, Kapteyn \citep{kapteyn}, NOVAS \citep{novas},
PALpy \citep{pal}, PyAST \citep[a wrapper for AST, described in][]{ast},
PyTPM\footnote{\url{https://github.com/phn/pytpm}}, PyEphem \citep{pyephem},
and pySLALIB \citep[a Python wrapper for SLALIB, described in][]{slalib}.

@@

The \package{astropy.table} subpackage provides functionality for
representing and manipulating heterogeneous data. In some respects,
this is similar to \package{numpy} record arrays \citep{numpy} or
\package{pandas} \texttt{DataFrame} objects \citep{pandas}, but with modifications for
astronomical data. Most notably, tables from \package{astropy.table}
allow for table or column metadata and can handle
vectors/arrays or arbitrary objects (with suitable column-like characteristics)
as table entries. The subpackage was described in
detail in \cite{astropy}.  Thus, in what follows, we only summarize
key new features or updates to \package{astropy.table} since the
previous \astropy paper. These are support for grouped table
operations, table concatenation, and using array-valued
\package{astropy} objects as table columns.

A table can contain data that naturally form groups; for example, it may
contain multiple observations of a few sources at different points in time
and in different bands. Then, we may want to split the table into groups based
on the combination of source observed and the band, after which we combine the
results for each combination of source and band in some way (e.g., finding
the mean or standard deviation of the fluxes or magnitudes over time) or filter
the groups based on user-defined criteria. These kinds of grouping and
aggregation operations are now fully supported by \texttt{Table} objects.

\texttt{Table} objects can now be combined in several different ways. If two
tables have the same columns, we may want to stack them ``vertically'' to create a
new table with the same columns but all rows. If two tables are row-matched but
have distinct columns, we may want to stack them ``horizontally'' to create a
new table with the same rows but all columns. For other situations, more generic
table concatenation or join are also possible when two tables share some
columns.

The \texttt{Table} object now allows array-valued \texttt{Quantity}, celestial
coordinate (\texttt{SkyCoord}), and date/time (\texttt{Time}) objects to
be used as columns. It also provides a general way for other user-defined
array-like objects to be used as columns.
This makes it possible, for instance, to easily
represent catalogs of sources or time series in \astropy, while having both the
benefits of the \texttt{Table} object (e.g., accessing specific rows/columns
or groups of them and combining tables) and of, for example,
the \texttt{SkyCoord} or the \texttt{Time} classes (e.g., converting the
coordinates to a different frame or accessing the date/time in the desired time scale).

Additionally, there is now interoperability between \texttt{Table}
and \package{pandas} \texttt{DataFrame} objects using the \texttt{Table.to_pandas()}
and \texttt{Table.from_pandas()} methods. \texttt{Table.to_pandas()} is,
however, limited to a subset of possible \texttt{Table} objects, because several
of the features outlined above cannot be represented as \package{pandas} data
frames. While these features might be contributed to or combined with
\package{pandas} at a later date, the architectural differences that
enable these areas are significant, so no one has yet volunteered to do this
major undertaking.

@@

The \package{astropy.io.fits} subpackage started as a direct port of the
PyFITS project \citep{PyFITS}. Therefore, it is pretty stable, with mostly bug
fixes but also a few new features and performance improvements.  The API
remains mostly compatible with PyFITS, which is now deprecated in favor of
\astropypkg.

Command-line scripts are now available for printing a summary of the HDUs in
FITS file(s) (\texttt{fitsinfo}) and for printing the header information to
the screen in a human-readable format (\texttt{fitsheader}).

FITS files are now loaded \emph{lazily} by default, i.e., an object
representing the list of HDUs is created but the data are not loaded into
memory until requested.  This approach should provide substantial speed-ups
when using the convenience functions (e.g., \texttt{getheader()} or
\texttt{getdata()}) to get an HDU that is near the beginning in a file with
many HDUs.

@@

The \package{astropy.modeling} subpackage provides a framework for representing
analytical models and performing model evaluation and parameter fitting.
The main motivation for this functionality was to create a framework that
allows arbitrary combination of models to support the Generalized World
Coordinate System (GWCS) package.\footnote{\url{https://github.com/spacetelescope/gwcs}}
The current FITS WCS specification lacks the flexibility to represent
arbitrary distortions and does not meet the needs of many types of current instrumentation.
The fact that the \astropypkg modeling framework now supports propagating
units also makes it a useful tool for representing and fitting astrophysical
models within data analysis tools.

@@

\package{astropy.modeling} provides several fitters which are wrappers around some of the \texttt{numpy} and \texttt{scipy.optimize} functions and provide support for specifying parameter constraints. The fitters take a model and data as input and return a copy of the model with the optimized parameter values set. The goal is to make it easy to extend the fitting framework to create new fitters. The optimizers available in \astropypkg version 2.0 are Levenberg--Marquardt (\texttt{scipy.optimize.leastsq}), Simplex (\texttt{scipy.optimize.fmin}), SLSQP (\texttt{scipy.optimize.slsqp}), and LinearLSQFitter (\texttt{numpy.linalg.lstsq} which provides exact solutions for linear models).

Modeling also supports a plugin system for fitters, which allows using the
\astropypkg models with external fitters. An example of this is
\package{SABA}\footnote{\url{https://github.com/astropy/saba}}, which is a bridge between
Sherpa \citep{sherpa},
and \package{astropy.modeling}, to bring the Sherpa fitters into \astropypkg.

\subsubsection{Creating New Models}
If arithmetic combinations of existing models is not sufficient, new model
classes can be defined in different ways. The \package{astropy.modeling}
package provides tools to turn a simple function into a full-featured model,
but it also allows extending the built-in model class with arbitrary code.

@@

The \package{astropy.modeling} subpackage now supports the representation,
evaluation, and fitting of models using \texttt{Quantity} objects, which attach
units to scalar values or arrays of values. In practice, this means that one
can, for example, fit a model to data with units and get parameters that also
have units out, or initialize a model with parameters with units and evaluate
it using input values with different but equivalent units. For example, the
blackbody model (\texttt{BlackBody1D}) can be used to fit observed flux
densities in a variety of units and as a function of different units of
spectral coordinates (e.g., wavelength or frequency).

\subsection{Convolution}
\label{ssec:convolution}

The \package{astropy.convolution} subpackage implements
\textit{normalized convolution} \citep[e.g.,][]{Knutsson1993}, an image
reconstruction technique in which missing data are ignored during the
convolution and replaced with values interpolated using the kernel.
An example is given in \figurename~\ref{fig:convolution-example}.
In \astropypkg versions $\leq 1.3$, the direct convolution and Fast Fourier Transform (FFT)
convolution approaches were inconsistent, with only the latter implementing
normalized convolution.
As of version 2.0, the two methods now agree and include a suite of
consistency checks.

@@

The \package{astropy.visualization} subpackage provides functionality that can
be helpful when visualizing data. This includes a framework (previously the
standalone \package{wcsaxes} package) for plotting
astronomical images with coordinates with \package{matplotlib}, functionality related to image
normalization (including both scaling and stretching), smart histogram
plotting, red-green-blue (RGB) color image creation from separate images,
and custom plotting styles for \package{matplotlib}.

@@

\package{matplotlib} allows a custom normalization and stretch to be used when
displaying data by passing in a normalization object.
The \package{astropy.visualization} package also provides a normalization class
that wraps the interval and stretches objects into a normalization object that
\package{matplotlib} understands.

@@

Astronomers dealing with observational imaging commonly need to make figures
with images that include the correct coordinates and optionally display a
coordinate grid. The challenge, however, is that the conceptual coordinate axes
(such as longitude/latitude) need not be lined up with the pixel axes of the
image. The \package{astropy.visualization.wcsaxes} subpackage implements a
generalized way of making figures from an image array and a WCS object
that provides the transformation between pixel and world coordinates.

@@

This subpackage differs from \package{APLpy} \citep{aplpy}, in that the latter
focuses on providing a very high-level interface to plotting that requires very
few lines of code to get a good result, whereas \package{wcsaxes} defines an
interface that is much closer to that of \package{matplotlib} \citep{matplotlib}.
This enables significantly more advanced visualizations.

An example of a visualization made with \package{wcsaxes} is shown in
\figurename~\ref{fig:wcsaxes}. This example illustrates the ability to
overlay multiple coordinate systems and customize which ticks/labels are shown
on which axes around the image. This also uses the image stretching
functionality from \sectionname~\ref{sec:stretch} to show the image in a
square-root stretch (automatically updating the tick positions in the colorbar).

\begin{figure}
\includegraphics[width=\textwidth]{cygnus_x_spitzer.pdf}
\caption{%
An example of figure made using the \package{astropy.visualization.wcsaxes}
subpackage, using \textit{Spitzer}/IRAC 8.0~$\mu$m data from the Cygnus-X
\textit{Spitzer} Legacy survey \citep{cygnusx}.
\label{fig:wcsaxes}
}
\end{figure}

@@

\package{astropy.visualization} also provides a histogram function, which is a
generalization of \texttt{matplotlib}’s histogram function, to allow for a more
flexible specification of histogram bins.  The function provides several methods
of automatically tuning the histogram bin size. It has a syntax identical to
\texttt{matplotlib}’s histogram function, with the exception of the \texttt{bins}
parameter, which allows specification of one of four different methods for
automatic bin selection: ``blocks'', ``knuth'', ``scott'', or ``freedman''.

@@

\cite{Lupton2004} describe an ``optimal'' algorithm for producing
RGB composite images from three separate high-dynamic range
arrays. The \package{astropy.visualization} subpackage provides a convenience
function to create such a color image.  It also includes an associated set of
classes to provide alternate scalings.
This functionality was contributed by developers from the Large Synoptic Survey
Telescope (LSST) and serves as an example of contribution to \astropy from a
more traditional engineering organization \citep{Jennes2016}.

The Sloan Digital Sky Survey (SDSS) SkyServer color images were made using a
variation on this technique.  As an example, in \figurename~\ref{fig:ngc6977},
we show an RGB color image of the Hickson 88 group, centered near
NGC~6977.\footnote{\url{http://skyserver.sdss.org/dr13/en/tools/chart/navi.aspx?ra=313.12381&dec=-5.74611}}
This image was generated from SDSS images using the
\package{astropy.visualization} tools.

>>>>

\begin{abstract}
The \astropy project supports and fosters the development of open-source and openly-developed
\python packages that provide commonly-needed functionality to the astronomical
community.
A key element of the \astropy project is the core package \astropypkg, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate and support
development for a broader ecosystem of inter-operable packages.
We conclude with a future outlook of planned new features and directions for the
broader \astropy project.
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
All modern astronomical research makes use of software in some way.
Astronomy as a field has thus long supported the development of software tools
for astronomical tasks, such as scripts that enable individual scientific
research, software packages for small collaborations, and data reduction
pipelines for survey operations.
Some software packages are, or were, supported by large institutions and are
intended for a wide range of users.
These packages therefore typically provide some level of documentation and user
support or training.
Other packages are developed by individual researchers or research groups and
are then typically used by smaller groups for more domain-specific purposes.
For both packages meant for wider distribution and for scripts specific to
particular research projects, a library that addresses common astronomical tasks
simplifies the software development process.
The users of such a library then also benefit from a community and ecosystem
built around a shared foundation.
The \astropy project has grown to become this community for \python astronomy
software, and the \astropypkg core package is a feature-rich \python library.

@@

The \astropy project aims to provide an open-source and open-development core
package (\astropypkg) and an ecosystem of \emph{affiliated packages} that
support astronomical functionality in the \python programming language.
The \astropypkg core package is now a feature-rich library of sufficiently
general tools and classes that supports the development of more specialized
code. An example of such functionality is reading and writing FITS files: It would be
time consuming and impractical for multiple groups to implement the FITS
standard \citep{FITS} and maintain software for such a general-purpose need.
Another example of such a common task is in dealing with representations of and
transformations between astronomical coordinate systems.

The \astropy project aims to develop and provide high-quality code and
documentation according to the best practices in software development.
The project makes use of different tools and web services to reach those
goals without central institutional oversight.
The first public release of the \astropypkg package is described in
\cite{astropy}. Since then, the \astropypkg package has been
used in hundreds of projects and the scope of the package has grown
considerably. At the same time, the scientific community
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropypkg core has
developed.
In this paper, we describe the current status of the \astropy community and the
\astropypkg core package and discuss goals for future development.

We start by describing the way the \astropy project functions and is organized
in \sectionname~\ref{sec:org}. We then describe the main software efforts
developed by the \astropy project itself: a core package called \astropypkg
(\sectionname~\ref{sec:core}) and several separate packages that help maintain
the infrastructure for testing and documentation
(\sectionname~\ref{sec:infrastructure}). We end with a short vision for
the future of \astropy and astronomical software in general in
\sectionname~\ref{sec:future}. The full paper, including the code
to produce the figures, is available in a \github
repository\footnote{\url{https://github.com/astropy/astropy-v2.0-paper}}

@@

A major part of the \astropy project is the concept of ``Affiliated Packages.''
An affiliated package is an astronomy-related
\python package that is not part of the \astropypkg core package, but
has requested to be included as part of the \astropy project's
community. These packages support the goals and vision of \astropy of
improving code re-use, interoperability, and embracing good coding
practices such as testing and thorough documentation.

Affiliated packages contain functionality that is more specialized,
have license incompatibilities, or have external dependencies (e.g., GUI
libraries) that make these packages more suitable to be separate from the
\astropypkg core package.
Affiliated packages may also be used to develop substantial new functionality
that will eventually be incorporated into the \astropypkg core package
(e.g., \astropywcsaxes).
New functionality benefits from having a rapid development and release cycle that is not tied to that of the \astropypkg core (\sectionname~\ref{sect:releasecycle}).

Affiliated packages are listed on the main \astropy website and advertised to
the community through \astropy mailing lists; a list of current affiliated
packages is included in \tablename~\ref{tab:registry}.
Becoming an affiliated package is a good way for new and existing packages to
gain exposure while promoting \astropy's high standard for code and
documentation quality.
This process of listing and promoting affiliated packages is one way in which
the \astropy project tries to increase code re-use in the astronomical
community.

@@

The \astropy project, as of the version 2.0 release, does not receive any direct
financial support for the development of \astropypkg.
Development of the software, all supporting materials, and community support are
provided by individuals who work on the \astropy project in their own personal
time, by individuals or groups contributing to \astropy as part of a research
project, or contributions from institutions that allocate people to work on
\astropy.
A list of organizations that have contributed to \astropy in this manner
can be found in the Acknowledgments.

Different funding models have been proposed for support of \astropy
(e.g., \citealt{2016arXiv161003159M}), but a long-term plan
for sustainability has not yet been established.
The \astropy project has the ability to accept financial contributions
from institutions or individuals through the NumFOCUS\footnote{NumFOCUS
is a 501(c)(3) nonprofit that supports and promotes world-class, innovative,
open source scientific computing.} organization. NumFOCUS has, to date, covered the direct costs incurred by the \astropy project.

\section{Astropy Core Package Version 2.0}
\label{sec:core}
The \astropy project aims to provide \python-based packages for all tasks that
are commonly needed in a large subset of the astronomical community.
At the foundation is the \astropypkg core package, which provides general
functionality (e.g., coordinate transformations, reading and writing
astronomical files, and units) or base classes for other
packages to utilize for a common interface (e.g.,
\href{http://docs.astropy.org/en/stable/nddata/index.html}{\texttt{NDData}}).
In this section, we highlight new features introduced or substantially improved
since version 0.2 (previously described in \citealt{astropy}).  The \astropypkg
package
provides a full log of changes\footnote{\url{https://github.com/astropy/astropy/blob/stable/CHANGES.rst}}
over the course of the entire project and more details about individual
subpackages are available in the documentation.\footnote{\url{http://docs.astropy.org/en/stable/}}
Beyond what is mentioned below, most subpackages have seen improved performance
since the release of the version 0.2 package.

@@

In order to check the accuracy of the coordinate transformations in
\astropysubpkg{coordinates}, we have created a set of benchmarks that we use
to compare transformations between a set of coordinate frames for a number of
packages\footnote{\url{http://www.astropy.org/coordinates-benchmark/summary.html}}.
Since no package can be guaranteed to implement all transformations to
arbitrary precision and some transformations are sometimes subject to
interpretation of standards (in particular in the case of Galactic coordinates),
we do not designate any of the existing packages as the ``ground truth'' but
instead compare each tool to all other tools. The benchmarks are thus useful
beyond the \astropy project since they allow all of the tools to be compared to
all other tools. The tools included in the benchmark at the moment include the
\astropypkg core package, Kapteyn \citep{kapteyn}, NOVAS \citep{novas},
PALpy \citep{pal}, PyAST \citep[a wrapper for AST, described in][]{ast},
PyTPM\footnote{\url{https://github.com/phn/pytpm}}, PyEphem \citep{pyephem},
and pySLALIB \citep[a Python wrapper for SLALIB, described in][]{slalib}.

@@

The \astropysubpkg{table} subpackage provides functionality for
representing and manipulating heterogeneous data. In some respects,
this is similar to \package{numpy} record arrays \citep{numpy} or
\package{pandas} dataframes \citep{pandas} but with modifications for
astronomical data. Most notably, tables from \astropysubpkg{table}
allow for table or column metadata and can handle
vectors or arrays as table entries.
The subpackage was described in
detail in \cite{astropy}.  Thus, in what follows, we only summarize
key new features or updates to \astropysubpkg{table} since the
previous \astropy paper. These are support for grouped table
operations, table concatenation, and using array-valued
\package{astropy} objects as table columns.

A table can contain data that naturally form groups; for example, it may
contain multiple observations of a few sources at different points in time
and in different bands. Then, we may want to split the table into groups based
on the combination of source observed and the band, after which we combine the
results for each combination of source and band in some way (e.g., finding
the mean or standard deviation of the fluxes or magnitudes over time) or filter
the groups based on user-defined criteria. These kinds of grouping and
aggregation operations are now fully supported by \astropyTable objects.

\astropyTable objects can now be combined in several different ways. If two
tables have the same columns, we may want to stack them ``vertically'' to create a
new table with the same columns but all rows. If two tables are row-matched but
have distinct columns, we may want to stack them ``horizontally'' to create a
new table with the same rows but all columns. For other situations, more generic
table concatenation or join are also possible when two tables share some
columns.

The \astropyTable object now allows array-valued \astropyQuantity, celestial
coordinate (\astropyskycoord), and date/time (\astropyTime) objects to
be used as columns. It also provides a general way for other user-defined
array-like objects to be used as columns.
This makes it possible, for instance, to easily
represent catalogs of sources or time series in \astropy, while having both the
benefits of the \astropyTable object (e.g., accessing specific rows/columns
or groups of them and combining tables) and of, for example,
the \astropyskycoord or the \astropyTime classes (e.g., converting the
coordinates to a different frame or accessing the date/time in the desired time scale).

@@

The \astropyiosubpkg{fits} subpackage started as a direct port of the
PyFITS project \citep{PyFITS}. Therefore, it is pretty stable, with mostly bug
fixes but also a few new features and performance improvements.  The API
remains mostly compatible with PyFITS, which is now deprecated in favor of
\astropypkg.

Command-line scripts are now available for printing a summary of the HDUs in
FITS file(s)
(\href{http://docs.astropy.org/en/stable/io/fits/usage/scripts.html#module-astropy.io.fits.scripts.fitsinfo}{\texttt{fitsinfo}})
and for printing the header information to the screen in a human-readable format
(\href{http://docs.astropy.org/en/stable/io/fits/usage/scripts.html#module-astropy.io.fits.scripts.fitsheader}{\texttt{fitsheader}}).

FITS files are now loaded \emph{lazily} by default, i.e., an object
representing the list of HDUs is created but the data are not loaded into
memory until requested.  This approach should provide substantial speed-ups
when using the convenience functions (e.g., \href{http://docs.astropy.org/en/stable/io/fits/api/files.html#getheader}{\texttt{getheader()}} or
\href{http://docs.astropy.org/en/stable/io/fits/api/files.html#getdata}{\texttt{getdata()}}) to get an HDU that is near the beginning in a file with
many HDUs.

\subsubsection{HDF5}

The \package{astropy.io.misc.hdf5} subpackage provides support for binary
read and write access to files in the \emph{Hierarchical Data Format}
(HDF5), if the \package{h5py} package is installed. Astropy table I/O
is offered transparently through \texttt{Table.read()} and
\texttt{Table.write()}, analogously to the other auto-detected
formats. The keyword \texttt{path='group/subgroup/dataset'} specifies
the path to the data inside the file's hierarchical structure.

@@

The \astropysubpkg{modeling} subpackage provides a framework for representing
analytical models and performing model evaluation and parameter fitting.
The main motivation for this functionality was to create a framework that
allows arbitrary combination of models to support the Generalized World
Coordinate System (GWCS) package.\footnote{\url{https://github.com/spacetelescope/gwcs}}
The current FITS WCS specification lacks the flexibility to represent
arbitrary distortions and does not meet the needs of many types of current instrumentation.
The fact that the \astropypkg modeling framework now supports propagating
units also makes it a useful tool for representing and fitting astrophysical
models within data analysis tools.

@@

\astropysubpkg{modeling} provides several fitters which are wrappers around some of the \texttt{numpy} and \texttt{scipy.optimize} functions and provide support for specifying parameter constraints. The fitters take a model and data as input and return a copy of the model with the optimized parameter values set. The goal is to make it easy to extend the fitting framework to create new fitters. The optimizers available in \astropypkg version 2.0 are Levenberg--Marquardt (\texttt{scipy.optimize.leastsq}), Simplex (\texttt{scipy.optimize.fmin}), SLSQP (\texttt{scipy.optimize.slsqp}), and LinearLSQFitter (\texttt{numpy.linalg.lstsq} which provides exact solutions for linear models).

Modeling also supports a plugin system for fitters, which allows using the
\astropypkg models with external fitters. An example of this is
\package{SABA}\footnote{\url{https://github.com/astropy/saba}}, which is a bridge between
Sherpa \citep{sherpa},
and \astropysubpkg{modeling}, to bring the Sherpa fitters into \astropypkg.

\subsubsection{Creating New Models}
If arithmetic combinations of existing models is not sufficient, new model
classes can be defined in different ways. The \astropysubpkg{modeling}
package provides tools to turn a simple function into a full-featured model,
but it also allows extending the built-in model class with arbitrary code.

@@

The \astropysubpkg{modeling} subpackage now supports the representation,
evaluation, and fitting of models using \astropyQuantity objects, which attach
units to scalar values or arrays of values. In practice, this means that one
can, for example, fit a model to data with units and get parameters that also
have units out, or initialize a model with parameters with units and evaluate
it using input values with different but equivalent units. For example, the
blackbody model
(\href{http://docs.astropy.org/en/stable/api/astropy.modeling.blackbody.BlackBody1D.html}{\texttt{BlackBody1D}})
can be used to fit observed flux densities in a variety of units and as a
function of different units of spectral coordinates (e.g., wavelength or
frequency).

\subsection{Convolution}

The \astropysubpkg{convolution} subpackage implements
\textit{normalized convolution} \citep[e.g.,][]{Knutsson1993}, an image
reconstruction technique in which missing data are ignored during the
convolution and replaced with values interpolated using the kernel.
An example is given in \figurename~\ref{fig:convolution-example}.
In \astropypkg versions $\leq 1.3$, the direct convolution and Fast Fourier Transform (FFT)
convolution approaches were inconsistent, with only the latter implementing
normalized convolution.
As of version 2.0, the two methods now agree and include a suite of
consistency checks.

@@

The \astropysubpkg{visualization} subpackage provides functionality that can
be helpful when visualizing data. This includes a framework (previously the
standalone \astropywcsaxes package) for plotting
astronomical images with coordinates with \package{matplotlib}, functionality related to image
normalization (including both scaling and stretching), smart histogram
plotting, red-green-blue (RGB) color image creation from separate images,
and custom plotting styles for \package{matplotlib}.

@@

\package{matplotlib} allows a custom normalization and stretch to be used when
displaying data by passing in a normalization object.
The \astropysubpkg{visualization} package also provides a normalization class
that wraps the interval and stretches objects into a normalization object that
\package{matplotlib} understands.

@@

Astronomers dealing with observational imaging commonly need to make figures
with images that include the correct coordinates and optionally display a
coordinate grid. The challenge, however, is that the conceptual coordinate axes
(such as longitude/latitude) need not be lined up with the pixel axes of the
image. The \astropywcsaxes subpackage implements a
generalized way of making figures from an image array and a WCS object
that provides the transformation between pixel and world coordinates.

@@

This subpackage differs from \package{APLpy} \citep{aplpy}, in that the latter
focuses on providing a very high-level interface to plotting that requires very
few lines of code to get a good result, whereas \astropywcsaxes defines an
interface that is much closer to that of \package{matplotlib} \citep{matplotlib}.
This enables significantly more advanced visualizations.

An example of a visualization made with \astropywcsaxes is shown in
\figurename~\ref{fig:wcsaxes}. This example illustrates the ability to
overlay multiple coordinate systems and customize which ticks/labels are shown
on which axes around the image. This also uses the image stretching
functionality from \sectionname~\ref{sec:stretch} to show the image in a
square-root stretch (automatically updating the tick positions in the colorbar).

\begin{figure}
\includegraphics[width=\textwidth]{cygnus_x_spitzer.pdf}
\caption{%
An example of figure made using the \astropywcsaxes
subpackage, using \textit{Spitzer}/IRAC 8.0~$\mu$m data from the Cygnus-X
\textit{Spitzer} Legacy survey \citep{cygnusx}.
\label{fig:wcsaxes}
}
\end{figure}

@@

\astropysubpkg{visualization} also provides a histogram function, which is a
generalization of \texttt{matplotlib}’s histogram function, to allow for a more
flexible specification of histogram bins.  The function provides several methods
of automatically tuning the histogram bin size. It has a syntax identical to
\texttt{matplotlib}’s histogram function, with the exception of the \texttt{bins}
parameter, which allows specification of one of four different methods for
automatic bin selection: ``blocks'', ``knuth'', ``scott'', or ``freedman''.

@@

\cite{Lupton2004} describe an ``optimal'' algorithm for producing
RGB composite images from three separate high-dynamic range
arrays. The \astropysubpkg{visualization} subpackage provides a convenience
function to create such a color image.  It also includes an associated set of
classes to provide alternate scalings.
This functionality was contributed by developers from the Large Synoptic Survey
Telescope (LSST) and serves as an example of contribution to \astropy from a
more traditional engineering organization \citep{Jennes2016}.

The Sloan Digital Sky Survey (SDSS) SkyServer color images were made using a
variation on this technique.  As an example, in \figurename~\ref{fig:ngc6977},
we show an RGB color image of the Hickson 88 group, centered near
NGC~6977.\footnote{\url{http://skyserver.sdss.org/dr13/en/tools/chart/navi.aspx?ra=313.12381&dec=-5.74611}}
This image was generated from SDSS images using the
\astropysubpkg{visualization} tools.

\begin{figure}
\includegraphics[width=\textwidth]{ngc6977}
\caption{An RGB color image of the region near the Hickson 88 group
constructed from SDSS images and the \astropysubpkg{visualization}
tools.
This example uses \astropywcsaxes to display the
sky coordinate grid, and the
\href{http://docs.astropy.org/en/stable/api/astropy.visualization.make_lupton_rgb.html}{\texttt{make\_lupton\_rgb()}}
function to produce the RGB image from three SDSS filter images ($g$, $r$, $i$).
The left and right panel images show two different parameter choices for the
stretch and softening parameters (shown in the titles).
\label{fig:ngc6977}}
\end{figure}
